---
title: "Board Games and Individual Ratings"
author: "Crystal Zhang, Lauren Grove, Nate Ren"
date: "5-1-24"
format: pdf
---

# Set Up

# Data Exploration 

## Data Structure 
```{r}

str(Boardgame)

```
There are a total of **10523** observations, each representing a different board game.
The dataset consists of **22** variables, which includes the types integer, string, and numerical. 
The integer variables are: `game_id`, `max_players`, `max_playtime`, `min_age`, ` min_players`, `min_playtime`, `playing_time`, `year_published`, and `users_rated`.
The character variables are: `description`, `image`, `name`, `thumbnail`, `artist`, `category`, `compilation`,``designer`, `expansion`, `family`, `mechanic`, and `publisher`. 
The only numerical variable is `average_rating`. 
More information on variables can be seen in the `Variable List` 

## Data Summary 
```{r}

summary(Boardgame)

```
This is a basic summary of the dataset. 
The distribution of integer and numberical variables, such as `max_playtime` and `min_players` are demonstrated with the 5-number summary and the mean. 
For character variables, only length (which is the same as the number of observations), class, and mode (both of these show the data type) are present.
